---
title: "Cleaner"
author: "Alfonso Poire"
date: "9/30/2021"
output: html_document
---
This script acts as a cleaner for the data that we want to put in to the RPPA Pipeline 

THIS FILE IS BY NO MEANS A FOOL PROFF CLEAN FOR ALL VARIATIONS YOU MUST CHECK YOUR DATA GOING IN AND OUT 

We will need to add variations as differing data formats come in as it seems like there is not consistent standardization in naming 

```{r}
library(tidyverse)
library(dplyr)
```
#Read in data and inspect what it looks like 
```{r}
data<-read.csv("./Imputs/Complete_SMMART_BRCA.csv",header=FALSE)
head(data)
```


#This Section is to fix data if the column names are in the fllowing format
  -Gordon Mills (Dong Zhang)-261035 ect.
```{r}
data<-as.data.frame(t(data))
names(data) <- as.matrix(data[1, ])
data <- data[-1, ]
colnames(data)[1] <- "Sample"

#seperate out the TCGA Samples and the normal samples
dataTCGA<-data%>%filter(grepl('TCGA', Sample))
dataNoNTCGA<-data%>%filter(!grepl('TCGA', Sample))

#Now lets fix the non tcga ones
#THIS WiLL ALSO DROP DASHES
dataNoNTCGA$Sample<-gsub("[^0-9.]", "", dataNoNTCGA$Sample)

#Now we need to recombine 
data2<-rbind(dataNoNTCGA,dataTCGA)
#Write it out 
data2<-as.data.frame(t(data2))
names(data2) <- as.matrix(data2[1, ])
data2 <- data2[-1, ]
#you should remove the sample name from the CSV manually for consistancy
write.csv(data2,"./Imputs/Cleaned_Complete_SMMART_BRCA.csv")
```

