---
title: "Heat_Mapper"
author: "Alfonso Poire"
date: "6/24/2021"
output: html_document
---
This script prepares the heatmaps generated in the RPPA report file and it takes in lvl 4 proteins and patient data. 

#Package read in 
```{r, messages=FALSE}
library(tidyverse)
library(dplyr)
library(janitor)
library(ComplexHeatmap)
library(data.table)
library(magick)
```
#Data Read in 
```{r,message=False}
data<-read.csv("./Imputs/Cleaned_Complete_SMMART_BRCA.csv",header=FALSE)
#if the data is not in ling format we need to roate it to get the TCGA groups labled
data_t<-as.data.frame(t(data))
data<-data_t%>%row_to_names(row_number = 1)
colnames(data)[1] <- c("rn")
#read in out TCGA Key
TCGA_Key<-read_tsv("Imputs/TCGASubtype.20170308.tsv")
```
Nowe we split out dataset based on the SMART vs TCGA criteria
```{r}
smart<-data%>%filter(!grepl("TCGA-*",rn))
TCGA<-data%>%filter(grepl("TCGA-*",rn))

```
now we can add out catigotization column to the patient dataset 
```{r}
smart<-cbind(Subtype="SMMART",smart)
#Reorder for readability 
smart<-smart%>%relocate(rn,Subtype)
#lets clean the first col 
smart$rn<-as.numeric(smart$rn)
smart$rn<-as.character(smart$rn)
```

Now we need to find the Subtypes for all of out TCGA Data
```{r}
#first thing we need to do is to replace any of the . in the name with - because that matches our key
#TCGA<-TCGA %>% 
  #mutate(rn = str_replace(rn, ".", "-"))
#we need to do a bit of data manipulation to get this ready for query so we will make a bunch of hold columns that we will later delete 
TCGA<-TCGA%>% separate(rn,sep="-",into=c("hold1","hold2","hold3","hold4","hold5"),extra="merge",remove=FALSE)
TCGA$hold4 <- gsub('.{1}$', '',TCGA$hold4)
#we are naming theis column sampleID so that we can match it to the key set later
TCGA$sampleID<-paste(TCGA$hold1,TCGA$hold2,TCGA$hold3,TCGA$hold4,sep="-")
#now lets do a quick reorder before we drop stuff 
TCGA<-TCGA%>%relocate(rn,sampleID)%>%select(-hold1,-hold2,-hold3,-hold4,-hold5)
head(TCGA)
#We will also create a query column outside the dataset so we can query our other data wth a single column
query<-TCGA%>%select(sampleID)
```
Now we are ready to use our query set to get the samples we need 
```{r}
query_results<-merge(query,TCGA_Key,by="sampleID",all.x = TRUE)
#now drop all the things we dont need
query_results<-query_results%>%select(sampleID,Subtype_mRNA)
#now we can add this bak to our TCGA Data
TCGA<-merge(TCGA,query_results,by="sampleID")
#Now reorder the columns so we can have a better looking set
TCGA<-TCGA%>%relocate(rn,sampleID,Subtype_mRNA)

```

lets also pull out the matches that we dident find 
```{r}
TCGA_Missing<-TCGA[is.na(TCGA$Subtype_mRNA),]
head(TCGA_Missing)
#now we will drop the nas from our TCGA set
TCGA<-TCGA[!is.na(TCGA$Subtype_mRNA),]
```
Now we need to merge our sets together again
```{r}
#rename the subtype col and drop the extra col 
TCGA<-TCGA%>%rename("Subtype"="Subtype_mRNA")
TCGA<-TCGA[-(2)]

full_data<-rbind(smart,TCGA)
```


First lets do the small subsample set
if we want to change the proteins in the heatmap this is the section to do it to do this please edit the Proteins_for_heatmap file in the imput file and save it as Proteins_for_heatmap_copy.txt
```{r}
proteins<-read.csv("./Imputs/Proteins_for_heatmap_copy.txt")
proteins<-c(proteins$rn)
```


```{r}
#this is our error correction box where we will add any unruly names that need to be fixed
full_dat<-full_data%>%rename("1433BETA"="X1433BETA","4EBP1"="X4EBP1","4EBP1_pS65"="X4EBP1_pS65","53BP1"="X53BP1")
#data_hm<-full_data%>%select(rn,Subtype,proteins)
data_hm<-full_data
```
Now enter your selections for the samples you want to use in this box for biopsy 1 and 2 
```{r}
b1<-"303010"
b2<-"303164"
```
we need to do some manipulation before we can do our calculations 
```{r}
data_hm<-as.data.frame(t(data_hm))
data_hm<-data_hm%>%row_to_names(row_number = 1)
#select the rows we want and get rid of any nas in thoes rows before we z score 
#make a function for this 
completeFun <- function(data, desiredCols) {
  completeVec <- complete.cases(data[, desiredCols])
  return(data[completeVec, ])
}

#now we will feed it ithe cols we want and remove the nas
data_hm<-completeFun(data_hm,c(b1,b2))
#im going to get rid of the subtype here not sure if thats a good idea perhaps we could ad it back in later but it seems unnecessary 
data_hm<-data_hm[-1,]
```
now we can z_score across
```{r}
data_hm[] <- lapply(data_hm, function(x) as.numeric(as.character(x)))
z_score<-sapply(data_hm, function(data_hm) (data_hm-mean(data_hm))/sd(data_hm))
rownames(z_score)<-rownames(data_hm)
z_score<-as.data.frame(z_score)
z_score<-z_score%>%select(all_of(b1),all_of(b2))
```

```{r}
#Lests get our fold change in there to 
z_score$fold_change<-z_score[b2]-z_score[b1]
#Rename the fold change col to be just fold change no mumbers 
colnames(z_score)[3] <- c("Fold_change")
```

Now lets make out heat maps 
```{r}
#first order the fold chnge to get a nice gradient 
z_score<-z_score[order(z_score$Fold_change),]
data_for_heatmap<-as.matrix(z_score)

data_for_heatmap2 = data_for_heatmap
colnames(data_for_heatmap) <- c("Bx1", "Bx2","Fold_change")
#data_for_heatmap2[data_for_heatmap2=="303145"] <- "Bx1"
#data_for_heatmap2[data_for_heatmap2=="272962"] <- "Bx2"


#data_for_heatmap2 %>%
  #rename(
   # Bx1 = `3031451`,
    #Bx2 = `272962`
  #)
```
Full set
```{r}
jpeg("./Outputs/heatmap_full_b1_b2.jpg", width = 500, height = 5000)
Heatmap(data_for_heatmap, 
        cluster_rows = FALSE,
        cluster_columns = FALSE, 
        show_column_names = TRUE, 
        show_row_names = TRUE,
        name="x",
        column_title = "Full Heatmap")
dev.off()
oldheatmap <-image_read("./Outputs/heatmap_full_b1_b2.jpg")
image_rotate(oldheatmap, 90) %>% image_write("./Outputs/rot_fullheat_b1_b2.jpg")
```
Now we will take the top and bottom 30 only 
```{r}
top20<-z_score%>%slice_max(Fold_change,n=30)
bottom20<-z_score%>%slice_min(Fold_change,n=30)
bottom20<-bottom20%>%arrange(desc(bottom20$Fold_change))
subset<-rbind(top20,bottom20)
data_for_heatmap<-as.matrix(subset)

colnames(data_for_heatmap) <- c("Bx1", "Bx2","Fold_change")
```
```{r}
jpeg("./Outputs/heatmap_most_variable_b1_b2.jpg", width = 500, height = 1000)
Heatmap(data_for_heatmap, 
        cluster_rows = FALSE,
        cluster_columns = FALSE, 
        show_column_names = TRUE, 
        show_row_names = TRUE,
        name= "x",
        column_title = "Most Variable Proteins",
        column_title_gp = gpar(fontsize = 20, fontface = "bold"),
        column_names_gp = gpar(fontsize = 18),
        #col = circlize::colorRamp2(c(-10, 0, 10), c("turquoise", "white","maroon" ))
        )
dev.off()

```
#If you want single patient heatmaps run this part 
seperate into individual patients 
```{r}
zscore_b1<-z_score%>%select(b1)
data_for_heatmap_b1<-as.matrix(zscore_b1)
zscore_b2<-z_score%>%select(b2)
data_for_heatmap_b2<-as.matrix(zscore_b2)
```

Full set
```{r}
jpeg("./Outputs/heatmap_full_b1.jpg", width = 500, height = 5000)
Heatmap(data_for_heatmap_b1, 
        cluster_rows = TRUE,
        cluster_columns = FALSE, 
        show_column_names = TRUE, 
        show_row_names = TRUE,
        name="x",
        column_title = "Bx 1 Heatmap")
dev.off()
oldheatmap <-image_read("./Outputs/heatmap_full_b1.jpg")
image_rotate(oldheatmap, 90) %>% image_write("./Outputs/rot_fullheat_b1.jpg")

jpeg("./Outputs/heatmap_full_b2.jpg", width = 500, height = 5000)
Heatmap(data_for_heatmap_b2, 
        cluster_rows = TRUE,
        cluster_columns = FALSE, 
        show_column_names = TRUE, 
        show_row_names = TRUE,
        name="x",
        column_title = "Bx 2 Heatmap")
dev.off()
oldheatmap <-image_read("./Outputs/heatmap_full_b2.jpg")
image_rotate(oldheatmap, 90) %>% image_write("./Outputs/rot_fullheat_b2.jpg")

```
Now we will take the top and bottom 30 only 
```{r}
top20<-z_score%>%slice_max(Fold_change,n=30)
bottom20<-z_score%>%slice_min(Fold_change,n=30)
bottom20<-bottom20%>%arrange(desc(bottom20$Fold_change))
subset<-rbind(top20,bottom20)

subset_b1<-subset%>% select(b1)
subset_b1<-as.matrix(subset_b1)
subset_b2<-subset%>% select(b2)
subset_b2<-as.matrix(subset_b2)

```
```{r}
jpeg("./Outputs/heatmap_most_variable_b1.jpg", width = 500, height = 1000)
Heatmap(subset_b1,
        cluster_rows = TRUE,
        cluster_columns = FALSE, 
        show_column_names = TRUE, 
        show_row_names = TRUE,
        name= "x",
        column_title = "Bx 1 Most Variable",
        column_title_gp = gpar(fontsize = 20, fontface = "bold"),
        column_names_gp = gpar(fontsize = 18),
        #col = circlize::colorRamp2(c(-10, 0, 10), c("turquoise", "white","maroon" ))
        )
dev.off
oldheatmap <-image_read("./Outputs/heatmap_most_variable_b1.jpg")
image_rotate(oldheatmap, 90) %>% image_write("./Outputs/rot_most_variable_b1.jpg")

jpeg("./Outputs/heatmap_most_variable_b2.jpg", width = 500, height = 1000)
Heatmap(subset_b2,
        cluster_rows = TRUE,
        cluster_columns = FALSE, 
        show_column_names = TRUE, 
        show_row_names = TRUE,
        name= "x",
        column_title = "Bx 2 Most Variable",
        column_title_gp = gpar(fontsize = 20, fontface = "bold"),
        column_names_gp = gpar(fontsize = 18),
        #col = circlize::colorRamp2(c(-10, 0, 10), c("turquoise", "white","maroon" ))
        )
dev.off()
oldheatmap <-image_read("./Outputs/heatmap_most_variable_b2.jpg")
image_rotate(oldheatmap, 90) %>% image_write("./Outputs/rot_most_variable_b2.jpg")

```

#This section is currently not workig 
#full set heat map 
The reason this isent working is because i forgot to swap the values to numeric 


we are going to add a column to denote our samples of intrest
```{r}
Bx1 = b1
Bx2 = b2

full_data<-full_data %>%mutate(ID = if_else(rn == b1,"Bx1", ifelse(rn==b2,"Bx2","SMMART")))
full_data<-full_data[!is.na(full_data$rn),]
```

```{r}

data_full_hm <- full_data[,-1]

rownames(data_full_hm) <- full_data[,1]

data_full_hm<-data_full_hm%>% arrange(Subtype)
#Get rid of all rows containing more than 50% NA aVAlue this can be changed
data_full_hm<-data_full_hm[, which(colMeans(!is.na(data_full_hm)) > 0.1)]
```
Seperate into SMART and TCGA cohorts
```{r}
Full_SMMART<-data_full_hm%>%filter(Subtype=="SMMART")
Full_TCGA<-data_full_hm%>%filter(Subtype!="SMMART")
```



```{r}
removeSamples<-Full_SMMART %>% #pull off the other annotation columns
  select(-Subtype,-ID)
removeSamples[] <- sapply(removeSamples, as.numeric)
removeSamples<-removeSamples[, which(colMeans(!is.na(removeSamples)) > 0.1)]
removeSamples<-removeSamples %>% mutate_if(is.numeric, funs(ifelse(is.na(.), 0, .)))

#removeSamples<-removeSamples[, which(colMeans(!is.na(removeSamples)) > 0.1)]
matrix_forHeatMap_SMMART<-as.matrix(removeSamples) #convert to matrix

matrix_forHeatMap_SMMART<-t(matrix_forHeatMap_SMMART)

#row annotation 
#this set-up is specific for the ComplexHeatmap package

row_ha = HeatmapAnnotation(sampleID = Full_SMMART$ID,col = list(bar = c("Bx1" = "red", "Bx2" = "green", "SMMART" = "Black")))
```
Lets sepera
```{r}
jpeg("./Outputs/big_Smmart.png", width =  2000, height = 3500)
Heatmap(matrix_forHeatMap_SMMART, 
        cluster_rows =TRUE,
        cluster_columns = TRUE, 
        show_column_names = FALSE, 
        show_row_names = TRUE,
        name= "DMSO adjusted Log2 value",
        top_annotation = row_ha
        )
dev.off()
oldheatmap_SMMARTbig <-image_read("./Outputs/big_Smmart.png")
image_rotate(oldheatmap_SMMARTbig, 90) %>% image_write("./Outputs/rot_big_SMMART.png")
```
```{r}
removeSamples<-Full_TCGA %>% #pull off the other annotation columns
select(-Subtype,-ID)
removeSamples[] <- sapply(removeSamples, as.numeric)
removeSamples<-removeSamples[, which(colMeans(!is.na(removeSamples)) > 0.1)]
removeSamples<-removeSamples %>% mutate_if(is.numeric, funs(ifelse(is.na(.), 0, .)))

#removeSamples<-removeSamples[, which(colMeans(!is.na(removeSamples)) > 0.1)]
matrix_forHeatMap_TCGA<-as.matrix(removeSamples) #convert to matrix

matrix_forHeatMap_TCGA<-t(matrix_forHeatMap_TCGA)

#row annotation 
#this set-up is specific for the ComplexHeatmap package

row_ha = HeatmapAnnotation(sampleID = Full_TCGA$ID,col = list(bar = c("Bx1" = "red", "Bx2" = "green", "SMMART" = "Black")))
```
Lets sepera
```{r}
jpeg("./Outputs/big_TCGA.png", width =  2000, height = 4500)
Heatmap(matrix_forHeatMap_TCGA, 
        cluster_rows =TRUE,
        cluster_columns = TRUE, 
        show_column_names = FALSE, 
        show_row_names = TRUE,
        name= "DMSO adjusted Log2 value",
        top_annotation = row_ha,
        )
dev.off()
oldheatmap_TCGAbig <-image_read("./Outputs/big_TCGA.png")
image_rotate(oldheatmap_TCGAbig, 90) %>% image_write("./Outputs/rot_big_TCGA.png")
```

FUll heatmaps top 30 
```{r}

subset<-cbind(proteins = rownames(subset), subset)
rownames(subset) <- 1:nrow(subset)
```

```{r}
SMMART_60v<-as.data.frame(matrix_forHeatMap_SMMART)
SMMART_60v <- cbind(proteins = rownames(SMMART_60v), SMMART_60v)
rownames(SMMART_60v) <- 1:nrow(SMMART_60v)

SMMART_60v<-SMMART_60v%>%filter(proteins %in% subset$proteins)
hold <- SMMART_60v[-1]
row.names(hold) <- SMMART_60v$proteins
SMMART_60v<-hold

SMMART_60v<-as.matrix(SMMART_60v)

row_ha = HeatmapAnnotation(sampleID = Full_SMMART$ID,col = list(bar = c("b1" = "red", "SMMART" = "Black")))
```
```{r}
jpeg("./Outputs/big_SMMART_60v.png", width =  2000, height = 1500)
Heatmap(SMMART_60v, 
        cluster_rows =TRUE,
        cluster_columns = FALSE, 
        show_column_names = FALSE, 
        show_row_names = TRUE,
        name= "DMSO adjusted Log2 value",
        top_annotation = row_ha,
        )
dev.off()
oldheatmap_SMMART60vbig <-image_read("./Outputs/big_SMMART_60v.png")
image_rotate(oldheatmap_TCGAbig, 90) %>% image_write("./Outputs/rot_big_SMMART60v.png")
```

```{r}

TCGA_60v<-as.data.frame(matrix_forHeatMap_TCGA)
TCGA_60v<- cbind(proteins = rownames(TCGA_60v), TCGA_60v)
rownames(TCGA_60v) <- 1:nrow(TCGA_60v)

TCGA_60v<-TCGA_60v%>%filter(proteins %in% subset$proteins)
hold <- TCGA_60v[-1]
row.names(hold) <- TCGA_60v$proteins
TCGA_60v<-hold

TCGA_60v<-as.matrix(TCGA_60v)

row_ha = HeatmapAnnotation(sampleID = Full_TCGA$ID,col = list(bar = c("b1" = "red", "b2" = "green", "SMMART" = "Black")))
```

```{r}
jpeg("./Outputs/big_TCGA60v.png", width =  2000, height = 1500)
Heatmap(TCGA_60v, 
        cluster_rows =TRUE,
        cluster_columns = TRUE, 
        show_column_names = FALSE, 
        show_row_names = TRUE,
        name= "DMSO adjusted Log2 value",
        top_annotation = row_ha,
        )
dev.off()
oldheatmap_TCGA60vbig <-image_read("./Outputs/big_TCGA60v.png")
image_rotate(oldheatmap_TCGAbig, 90) %>% image_write("./Outputs/rot_big_TCGA60v.png")
```