---
title: "patway_prepare"
author: "Alfonso Poire"
date: "6/4/2021"
output: html_document
---
This script will take in a pathway score file and out_put a format that can be passed up to the RPPA Report generator script

Is this suppose to be the one from before I think it is 

Packages
```{r setup,echo=FALSE}
library(tidyverse)
library(dplyr)
library(stringr)
library(grid)
```

data read in 
Change the data pathways to the file your using 
```{r,echo=FALSE,message=FALSE}
data<-read.csv("Outputs/Cleaned_Complete_SMMART_BRCA_pathway_score_data.csv",header=TRUE)
data<-as.data.frame(t(data))
names(data) <- data[1, ]
data <- data[-1, ]
data[is.na(data)] <- 0
#Rename the first column so that we can manipulate it more easily
colnames(data)[1]<-"Sample"
TCGA_Key<-read_tsv("Imputs/TCGASubtype.20170308.tsv")
```

now lets split the Data Set into 2 separate categories the TCGA data and the patients 
lets clan the names before we do anything else 
```{r}
#This will get rid of the Score part we could also change it to Pathway* and that would get rid of everything but the name
for (col in 1:ncol(data)){
  colnames(data)[col]<-sub("_Pathway_Score*", "", colnames(data)[col])
}

```

```{r}
patients<-data%>%filter(!grepl("TCGA-*",Sample))
TCGA<-data%>%filter(grepl("TCGA-*",Sample))
```

now we can add out categorization column to the patient Data Set 
```{r}
patients<-cbind(Subtype="SMMART",patients)
#Reorder for readability 
patients<-patients%>%relocate(Sample,Subtype)
patients$Sample<-as.numeric(patients$Sample)
patients$Sample<-as.character(patients$Sample)
```

Now we need to find the  for all of out TCGA Data
```{r}
#first we need to do a bit of data manipulation to get this ready for query so we will make a bunch of hold columns that we will later delete 
TCGA<-TCGA%>% separate(Sample,sep="-",into=c("hold1","hold2","hold3","hold4","hold5"),extra="merge",remove=FALSE)
TCGA$hold4 <- gsub('.{1}$', '',TCGA$hold4)
#we are naming this column sampleD so that we can match it to the key set later
TCGA$sampleID<-paste(TCGA$hold1,TCGA$hold2,TCGA$hold3,TCGA$hold4,sep="-")
#now lets do a quick reorder before we drop stuff 
TCGA<-TCGA%>%relocate(Sample,sampleID)%>%select(-hold1,-hold2,-hold3,-hold4,-hold5)
head(TCGA)
#We will also create a query column outside the Data Set so we can query our other data with a single column
query<-TCGA%>%select(sampleID)
```

Now we are ready to use our query set to get the samples we need 
```{r}
query_results<-merge(query,TCGA_Key,by="sampleID",all.x = TRUE)
#now drop all the things we don't need
query_results<-query_results%>%select(sampleID,Subtype_mRNA)
#now we can add this back to our TCGA Data
TCGA<-merge(TCGA,query_results,by="sampleID")
#Now reorder the columns so we can have a better looking set
TCGA<-TCGA%>%relocate(Sample,sampleID,Subtype_mRNA)

```
lets also pull out the matches that we did not find 
```{r}
TCGA_Missing<-TCGA[is.na(TCGA$Subtype_mRNA),]
head(TCGA_Missing)
#now we will drop the nas from our TCGA set
TCGA<-TCGA%>%drop_na()
```

#now if we want we can select a secific patitent if we would like to from our patient data set heres where we do it 
```{r}
#Print out the available samples
unique(patients$Sample)
```



#FILL THIS LIST WITH THE SAMPLES YOU WISH TO SELECT
```{r}
choices<-c("303010","303164")

selected_patient<-patients%>%filter(Sample %in% choices)
```
#We can add a density map for that patient here
```{r}
long_data<-selected_patient%>%pivot_longer(cols=-(1:2),names_to = "Pathway")
line_data<-long_data %>%select(-Subtype)%>%filter(Pathway!="Pathway")
line_data$value<-as.numeric(line_data$value)

line_data_Bxrename = line_data
line_data_Bxrename[line_data_Bxrename=="303010"] <- "Bx1"
line_data_Bxrename[line_data_Bxrename=="303164"] <- "Bx2"
```

#First lets get the TCGA Graphs made
```{r}
TCGA_Graph_Data<-TCGA%>%select(-Sample,-sampleID,-Pathway)
#rotate it longer
#pivot longer
TCGA_long_data<-pivot_longer(TCGA_Graph_Data,cols=-(1),names_to = "Pathway")
TCGA_long_data$value<-as.numeric(TCGA_long_data$value)

```
```{r}
#We want to remove something choose it here
#the only note here is that once removed the whole file must be run again to replace
# "Her2"   "LumA"   "LumB"   "Basal"  "Normal"
#TCGA_long_data<-TCGA_long_data%>%filter(Subtype_mRNA!="Her2")
#TCGA_long_data<-TCGA_long_data%>%filter(Subtype_mRNA!="LumA")
#TCGA_long_data<-TCGA_long_data%>%filter(Subtype_mRNA!="LumB")
#TCGA_long_data<-TCGA_long_data%>%filter(Subtype_mRNA!="Basal")
TCGA_long_data<-TCGA_long_data%>%filter(Subtype_mRNA!="Normal")
```
```{r}
#IF you want to change the colors for the samples change this section
vars=c('Her2'='purple','LumA'='Green','LumB'='Blue','Basal'='orange','Normal'='red')
```

```{r}
#Not ure why the legend isent showing up 
TCGA_histograms <- ggplot(TCGA_long_data, aes(value,group=Subtype_mRNA,fill=Subtype_mRNA)) +
  #Change
  geom_density(adjust=1.5, alpha=.4)+
  theme_minimal()+
  #THis part needs to be eddited to the patient of interest
  ggtitle("Density Maps for TCGA Tumors, Patient XXX Highlighted")+
  xlab('Value')+ylab('density') + 
  facet_wrap(~Pathway)+  
  geom_vline(data =line_data_Bxrename, aes(xintercept = value,color=Sample),size = 3, alpha=.4)

TCGA_histograms<-TCGA_histograms+
  theme(strip.text.x = element_text(size=30),plot.title = element_text(size=70))+
  theme(axis.text=element_text(size=30),axis.title=element_text(size=40,face="bold"))+
  #scale_color_manual(name = 'Subtype mRNA',values=vars)+
  theme(legend.text=element_text(size=30),legend.title =element_text(size=40))

ggsave("./Outputs/TCGA_histograms.jpeg",TCGA_histograms,width=40,height=40)
```

```{r}
#Uncomment this if you want to save each graph individually

plot_list = list()
loop.list <- unique(TCGA_long_data$Pathway)
for (i in seq_along(loop.list)) {
  title_name=paste(i)
  lines_hold<-line_data_Bxrename%>%filter(Pathway==loop.list[i])
  p = ggplot(TCGA_long_data[TCGA_long_data$Pathway == loop.list[i], ],
             aes(value,group=Subtype_mRNA,fill=Subtype_mRNA)) +
    geom_density(adjust=1.5, alpha=.4)+
    theme_minimal()+
    ggtitle(loop.list[i])+
    theme(plot.title = element_text(hjust = 0.5))+
    #theme(axis.text=element_text(size=30),axis.title=element_text(size=40,face="bold"))+ if you want to change the size of the title or the axis titles   uncomment this 
  # theme(legend.text=element_text(size=30),legend.title =element_text(size=40))+
    xlab('Value')+ylab('density') +
    geom_vline(data =lines_hold, aes(xintercept = value,color=Sample),size = 2)#if the lines are stacked on top of  add this , alpha=.4

  plot_list[[i]] <- ggplotGrob(p)
}
#seq_along(loop.list)
for (i in seq_along(loop.list)) {
  file_name = paste("./Outputs/Density_plot_",loop.list[i], ".jpeg", sep="")
  tiff(file_name)
  grid.draw(plot_list[[i]])
  dev.off()
}
```

#Lets try and do smart
```{r}
patients<-patients%>%select(-Subtype,-Pathway)


patients_graph_data<-pivot_longer(patients,cols=-(1),names_to = "Pathway")
patients_graph_data$value<-as.numeric(patients_graph_data$value)
```



```{r}
SMART_histograms <- ggplot(patients_graph_data, aes(value)) +
  geom_density(fill="turquoise",adjust=1.5, alpha=.4) + 
  #theme(legend.position = "none")+
  theme_minimal()+
  facet_wrap(~ Pathway)+
  ggtitle("Density Maps for SMMART Tumors, Patient XXX Highlighted")+
  geom_vline(data =line_data_Bxrename, aes(xintercept = value,color=Sample), size = 3)

SMART_histograms<-SMART_histograms+
  theme(strip.text.x = element_text(size=30),plot.title = element_text(size=70))+
  theme(axis.text=element_text(size=30),axis.title=element_text(size=50,face="bold"))+
  theme(legend.text=element_text(size=30),legend.title =element_text(size=40))

ggsave("./Outputs/SMART_histograms.jpeg",SMART_histograms,width=40,height=40)
```
```{r}
# # this if you want to save each graph individually
# 
# plot_list = list()
# loop.list <- unique(patients_graph_data$Pathway)
# for (i in seq_along(loop.list)) {
#   title_name=paste(i)
#   lines_hold<-line_data%>%filter(Pathway==loop.list[i])
#   p = ggplot(patients_graph_data[patients_graph_data$Pathway == loop.list[i], ], 
#              aes(value)) +
#     geom_density(fill="turquoise",adjust=1.5, alpha=.4)+
#     ggtitle(loop.list[i])+
#     geom_vline(data =lines_hold, aes(xintercept = value,color=Sample))
#   
#   plot_list[[i]] <- ggplotGrob(p)
# }
# 
# for (i in seq_along(loop.list)) {
#   file_name = paste("./Outputs/Density_plot_",loop.list[i], ".png", sep="")
#   tiff(file_name)
#   grid.draw(plot_list[[i]])
#   dev.off()
# }

```

