---
title: "Score_Calculator"
author: "Alfonso Poire"
date: "6/11/2021"
output: html_document
---
THis script cleans and rotates the data from the Pathway_Score_calculator.ipynb file and is run half way through it then the rest of the other script must be run to complete pathway scoring 

#We are going to take the output from the python file and nest it by the pathway and we will then do the removals we want to do 
then we can put it out again 

```{r,echo=False}
library(tidyverse)
library(dplyr)
```

Data read in 
This data must be changed to the correct file that you are working with ex ______Score_pass.csv
```{r}
  data<-as.data.frame(t(read.csv("./Outputs/SMMART_EBN_RBN_5_score_pass.csv",header=FALSE)))
pathways<-read.csv("./Imputs/Pathway_Scores.txt",sep="\t")
```
now set the header to the patient id renaming the rows is a waste of time 
```{r}
names(data)<-data[1,]
data<-data[-1,]
#we also need to rename the first column so we can merge on it 
colnames(data)[1]<-c("Predictor")
```
Now lets prepare the pathways set
```{r}
#first lets get rid of the cols from pathway score we dont need 
pathways<-pathways%>%select(-Weight,-Count)

#lets clean the predictor column so it matches out format
#ASK ABOUT BLC2
pathways$Predictor<-str_replace_all(pathways$Predictor, "[^[:alnum:]]", "")
pathways$Predictor<-toupper(pathways$Predictor)

```

now we are going to create a pathway column by using the pathway score as a dictionary 
```{r}
#now lets do the merge
data<-merge(data,pathways,by="Predictor")
write.csv(data,"./Outputs/Score_pass_back.csv",row.names=FALSE)
```

#Reformating section this needs to be somewhere better 
```{r}
# data<-read.csv("./Imputs/adj_soonyoung.csv")
# 
# data_test<-data %>% 
#   group_by(Sample_description) %>% summarise_all(funs(mean(., na.rm = TRUE)))
# 
# data<-data_test%>%select(-X,-Cell_line,-Exparement,-Antibody_Name)
# 
# data<-t(data)
# colnames(data) <- data[1,]
# data<- data[-1, ]
# data<- data[-1, ]
# write.csv(data,"non_adj_soonyoung_score_ready.csv")
```


